%!TEX root = NeedleReference.tex
\chapter{Database Testing}

When unit-testing your application, it is usually recommended to mock calls to DAOs oder database dependent services in your test cases.
However, you will also need to test against a real database, e.g. to make sure that your queries work as expected. 

\section{Database Testcase}

In these cases, Needle can automatically create and inject the \verb|EntityManager| instance into your objects under test. 
You simply need to provide a JPA persistence.xml and the JDBC driver to the classpath of the test execution. 
The \verb|persistence.xml| file is the standard configuration file in JPA. The \verb|persistence.xml| file must define a persistence-unit with a unique name and the transaction type RESOURCE\_LOCAL for a Java SE environment. 

The following listing below shows a complete example of a \verb|persistence.xml| file. 

\begin{lstlisting}[caption=test persistence.xml]
<?xml version="1.0" encoding="UTF-8" ?>
<persistence xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
   xsi:schemaLocation="http://java.sun.com/xml/ns/persistence http://java.sun.com/xml/ns/persistence/persistence_2_0.xsd"
   version="2.0" xmlns="http://java.sun.com/xml/ns/persistence">
   <persistence-unit name="TestDataModel"
      transaction-type="RESOURCE_LOCAL">

      <class>de.akquinet.jbosscc.needle.example.User</class>
      <class>de.akquinet.jbosscc.needle.example.Profile</class>

      <properties>
         <property name="javax.persistence.jdbc.driver" value="org.hsqldb.jdbcDriver" />
         <property name="javax.persistence.jdbc.url" value="jdbc:hsqldb:." />
         <property name="javax.persistence.jdbc.user" value="sa" />
         <property name="javax.persistence.jdbc.password" value="" />

         <!-- EclipseLink should create the database schema automatically -->
         <property name="eclipselink.ddl-generation" value="create-tables" />
         <property name="eclipselink.ddl-generation.output-mode"
            value="database" />
      </properties>

   </persistence-unit>
</persistence>
\end{lstlisting}

The UserTest below checks the JPA mapping against a real database.

\begin{lstlisting}[language={JAVA},caption=User persistence test]
public class UserTest {
   @Rule
   public DatabaseRule databaseRule = new DatabaseRule();

   @Test
   public void testPersist() throws Exception {
      EntityManager entityManager = databaseRule.getEntityManager();

      User user = new UserTestdataBuilder(entityManager).buildAndSave();

      User userFromDb = entityManager.find(User.class, user.getId());
      Assert.assertEquals(user.getId(), userFromDb.getId());
   }
}
\end{lstlisting}

Note: The persistence context is cleared after each test!

\section{Transaction utilities}

The EntityManager is the primary interface used by application developers to interact with the underlying database.
Many operations must be executed in a transaction which is not started in the test component, because it is ususally maintained by
the application server. To run your code using transactions the \verb|TransactionHelper| Utility makes this more convenient.

\begin{lstlisting}[language={JAVA},caption=Transaction utilities]
public class UserTest {
   @Rule
   public DatabaseRule databaseRule = new DatabaseRule();
   
   private TransactionHelper transactionHelper = databaseRule.getTransactionHelper();

   @Test
   public void testPersist() throws Exception {

      final User user = new User();
      
      ...
      
      transactionHelper.executeInTransaction(new VoidRunnable() {
         @Override
         public void doRun(EntityManager entityManager) throws Exception {
            entityManager.persist(user);
         }
      });
      
      ...
      
   }
}
\end{lstlisting}

The above example illustrates the use of the \verb|TransactionHelper| class and the execution of a transaction. The implementation of the VoidRunnable is executed within a transaction. There is also \verb|Runnable<T>| interface that contains a method run return a value to be used for further testing.

\section{Database operation}

A common issue in unit tests that access a real database is their effect on the state of the persistence store.
They usually expect a certain state at the beginning and alter it at runtime which of course will affect other tests.
To address that problem optional Database operations can be executed before and after test execution, in order to clean up or set up data. 

There are the following implementations:
\begin{enumerate}
\item \verb|de.akquinet.jbosscc.needle.db.operation.ExecuteScriptOperation| \\ Execute sql scripts during test setup and tear down. A before.sql and after.sql script must be provided on the classpath.
\item \verb|de.akquinet.jbosscc.needle.db.operation.h2.H2DeleteOperation| \\ Deletes all rows of all tables of the h2 database.
\item \verb|de.akquinet.jbosscc.needle.db.operation.hsql.HSQLDeleteOperation| \\ Deletes all rows of all tables of the hsql database.
\end{enumerate}

To use own Database operation implementations, the abstract base class \verb|de.akquinet.jbosscc.needle.db.operation.AbstractDBOperation| must be implemented and configured in the \verb|needle.properties| file.

\section{Testdatabuilder}

Using the Test Data Builder pattern, the builder class provides methods that can be used to configure the test objects. 
Properties that are not configured use default values. The builder methods can be chained together and provide transient or persistent testdata for the test case.

For this purpose Needle provides an abstract base class. The following code examples shows two implementations of Test Data Builder pattern. The Testdatabuilder inherit from 
de.akquinet.jbosscc.needle.db.testdata.\-AbstractTestdataBuilder class. 

\begin{lstlisting}[language={JAVA},caption=User Testdatabuilder]
public class UserTestdataBuilder extends AbstractTestdataBuilder<User> {

   private String withUsername;
   private String withPassword;
   private Profile withProfile;

   public UserTestdataBuilder() {
      super();
   }

   public UserTestdataBuilder(EntityManager entityManager) {
      super(entityManager);
   }

   public UserTestdataBuilder withUsername(final String username) {
      this.withUsername = username;
      return this;
   }

   private String getUsername() {
      return withUsername != null ? withUsername : "username";
   }

   public UserTestdataBuilder withPassword(final String password) {
      this.withPassword = password;
      return this;
   }

   private String getPassword() {
      return withPassword != null ? withPassword : "password";
   }

   public UserTestdataBuilder withProfile(final Profile profile) {
      this.withProfile = profile;
      return this;
   }

   private Profile getProfile() {
      if (withProfile != null) {
         return withProfile;
      }

      return hasEntityManager() ? new ProfileTestdataBuilder(
            getEntityManager()).buildAndSave()
            : new ProfileTestdataBuilder().build();

   }

   @Override
   public User build() {
      User user = new User();
      user.setUsername(getUsername());
      user.setPassword(getPassword());
      user.setProfile(getProfile());
      return user;
   }

}
\end{lstlisting}

\begin{lstlisting}[language={JAVA},caption=Profile Testdatabuilder]
public class ProfileTestdataBuilder extends AbstractTestdataBuilder<Profile> {

   private String withLanguage;
   
   public ProfileTestdataBuilder() {
      super();
   }

   public ProfileTestdataBuilder(EntityManager entityManager) {
      super(entityManager);
   }


   public ProfileTestdataBuilder withLanguage(final String language) {
      this.withLanguage = language;
      return this;
   }

   private String getLanguage() {
      return withLanguage != null ? withLanguage : "de";
   }

   @Override
   public Profile build() {
      Profile profile = new Profile();
      profile.setLanguage(getLanguage());
      return profile;
   }
}
\end{lstlisting}

In the example the Testdatabuilder is using defaults for everything except the username.

\begin{lstlisting}[language={JAVA},caption=Build an persistent User object]
final User user = new UserTestdataBuilder(databaseRule.getEntityManager()).withUsername("user").buildAndSave();
\end{lstlisting}

\begin{lstlisting}[language={JAVA},caption=Build an transient User object]
final User user = new UserTestdataBuilder().withUsername("user").build()
\end{lstlisting}

\section{Injectable database resources}

In combination with the NeedleRule, the following resources are injectable by the DatabaseTestcase.

\begin{enumerate}
\item EntityManager
\item EntityManagerFactory
\item EntityTransaction
\item TransactionHelper
\end{enumerate}

Example:

\begin{lstlisting}[language={JAVA},caption=Injectable database resources]
public class UserDaoTest{
    
    @Rule
    public NeedleRule needleRule = new NeedleRule().withOuter(new DatabaseRule());
    
   //  Alternative:
   // 
   //  @Rule
   //  public DatabaseRule dbRule = new DatabaseRule()
   //  @Rule
   //  public NeedleRule needleRule = new NeedleRule(dbRule);
   

    @Inject
    private EntityManager entityManager;

    @Inject
    private EntityTransaction entityTransaction;

    //…
}
\end{lstlisting}
